Supplementary Material for ``Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)"

### 2023-07-30 ba_resubmit/kundinger_supplement_final.tex
add supplement.

@@

%% *** Frontmatter *** 
\linenumbers
\begin{frontmatter}
\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
%\title{A Sample Document\thanksref{T1}}
%\thankstext{T1}{<thanks text>}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

@@

\begin{abstract} 
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, \textcolor{red}{unlike} the sequential updates of \texttt{BRL}.}
\textcolor{teal}{We ensure \textcolor{red}{\text{fabl} results in a bipartite matching} using decision theoretic post-processing techniques from the existing literature.}
\textcolor{red}{To reduce computational overhead and memory costs, we use hashing and a new technique that we call storage efficient indexing. }
%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead. Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing. 
%}
%This allows
%us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. 
Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

%% ** Keywords **
\begin{keyword}
\kwd{record linkage}
\kwd{entity resolution}
\kwd{data cleaning}
\kwd{parallelization}
\kwd{hashing}
\end{keyword}

@@

%% ** Mainmatter **

\section{Introduction}
\label{sec:introduction}

Before conducting data analysis, it is often necessary to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. In their approach, the data analyst first creates comparison vectors for each pair of records in the data files. These vectors indicate how similar the records are on a set of variables measured in both files, known as the linkage variables.  Using these comparison vectors, the analyst classifies each pair as a match or nonmatch using a likelihood ratio test. Crucially, these decisions are made independently for each pair. The \cite{fellegi_theory_1969} approach has been extended for a wide variety of applications \citep[e.g.,][]{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2022prior}. In this article, we build on the contributions to the comparison vector approach.

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}

Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. 
\textcolor{red}{We use independent prior distributions for the matching status of each record, an idea suggested in \cite{wortman2019}.}
%% Jerry thought the wording below was too vague and made a suggestion that is more direct and clear. 
%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} 
%
\textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, \textcolor{red}{unlike the} sequential updates of \cite{sadinle_bayesian_2017}. We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
This \textcolor{red}{approach} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study. Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.

\section{Review of Fellegi-Sunter Approaches for Record Linkage}
\label{sec:review-of_prior-work}

Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files. For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.  Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.  Without loss of generality, denote files such that $n_A \geq n_B$.

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.

When linking variable $f$ is categorical, a common way to define $\gamma_{ij}^f$ is an indicator for exact agreement. For example, if zip code is linking variable $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical linking variables, we can use the absolute difference of the two values. For example, if age is linking variable $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text variables like names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for linking variable $f$ of records $A_i$ and $B_j.$ The range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

%\textcolor{red}{In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. 
%We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.}

Having defined comparison vectors, we now turn to the \cite{fellegi_theory_1969} model for record linkage.  We begin with notation for defining linkages. Under bipartite matching, the set of matches across $A$ and $B$ can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_An_B}$, where
\begin{align}
\Delta_{ij} =
\begin{cases}
1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
0, & \text{otherwise}.\\
\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
\begin{align}
Z_{j} =
\begin{cases}
i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise. When presenting the models, we use both representations for convenience.

Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.  For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these independence assumptions,  one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.  We have
\begin{subequations}
\begin{align}
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  \textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
%
%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
%
It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior distribution used by \cite{sadinle_bayesian_2017} is given by
\begin{align}
\label{eqn:sadinle_prior}
p(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently \textcolor{teal}{sequential}, which can be slow when working on linkage tasks with more than a few thousand records.

\section{Fast Beta Linkage}
\label{sec:fast-beta-linkage}

\textcolor{red}{For \text{fabl}, we also use (\ref{eqn:fs_model}) and (\ref{eqn:m_dist}), along with Dirichlet prior distributions for the $\mathbf{m}$ and $\mathbf{u}$ parameters.  However, instead of the prior over $\bm{Z}$}
%Instead of the prior over $\bm{Z}$ 
from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, 
\textcolor{teal}{we propose an alternative, called the fast beta linkage (fabl) prior.}
%the fast Beta prior as follows. 
For each $Z_j$, we have

\begin{subequations}
\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
\end{align}
\end{subequations}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

\begin{rem}
\textcolor{red}{
In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, this highly informative prior weighting on matching can result in overly optimistic match rates. Hence, \textcolor{red}{we prefer (\ref{eqn:fast_beta_prior}).} We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
}
\end{rem}

\begin{rem}
Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in \textcolor{red}{the \cite{fellegi_theory_1969} model} estimates the proportion of record pairs that are matches. In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}). In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by \textcolor{red}{$1/n_A$} and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence. 
\end{rem}

\textcolor{red}{In many record linkage tasks, the records in $A$ or $B$ have missing fields, so that some $\gamma_{ij}^f$ could be missing.  To handle this, we follow the approach described in \cite{sadinle_bayesian_2017}.  Let $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. We assume that any missing comparisons are missing at random (MAR) or missing completely at random (MCAR). For likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}, so that we can marginalize over the missing data and perform all computation using the observed data \citep{sadinle_detecting_2014, sadinle_bayesian_2017}.}

%\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors. To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing at random (MAR) or missing completely at random (MCAR). It is important to note that for likelihood based inference, both MAR and MCAR lead to ignorability, or rather, being able to consider the likelihood for the observed data \citep[Section 6.2]{LittleRubin2002}. Therefore, we can marginalize over the missing data and perform all computation using the observed data, leading to (\ref{eqn:likelihood}). For further details on these connections, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.
%}

\textcolor{teal}{We summarize \textcolor{red}{the model for \text{fabl}} as follows:}

\begin{subequations}
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}),\quad  \forall  \; f = 1, \ldots, F, \label{eqn:m} \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}), \quad \forall \; f = 1, \ldots, F,  \label{eqn:u}\\
	p(Z_j = q| \pi)  &=
	\begin{cases} 
		\frac{1}{n_A}\pi,  & q \leq n_A; \\
		1-\pi, &  q  = n_A + j; \\
	\end{cases} \label{eqn:z}\\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi-prior},
\end{align}
\end{subequations}
\textcolor{teal}{where recall that \textcolor{red}{$(\alpha_{\pi}, \beta_{\pi})$} are fixed hyper-parameters.}

%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,

%We \textcolor{teal}{approximate} the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. 
%We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^{(s)}$, we use the full conditionals,

\textcolor{teal}{We \textcolor{red}{now} present a Gibbs sampler to approximate the joint posterior distribution of $\bm{Z}, \bm{m},$ $\bm{u},$ and $\pi$ given the observed comparison data from the likelihood and priors assumed in (\ref{eqn:likelihood} - \ref{eqn:pi-prior}). 
%While we could marginalize over $\pi$ and derive a collapsed Gibbs sampler, in the spirit of \cite{sadinle_bayesian_2017}, this leads to sequential updates, which are computationally slow. Instead, we proceed with a traditional Gibbs sampler, where the updates to the matching label (or coreference matrix) are parallel, allowing for more efficient updates.
Refer to Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}
%%
We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. \textcolor{blue}{Brian: state how you initialize $\pi$.} \textcolor{teal}{Denote the current draw of the sampler by $s$ and the updated draw by $s+1$.} \textcolor{red}{To update $\mathbf{m}$ and $\mathbf{u}$, for all, $f=1, \ldots, F,$ sample}

%\begin{subequations}
%\begin{align}
%	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  \\
%	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
%\end{align}
%\end{subequations}

%\begin{subequations}
%
%	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
%	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
%\end{align}
%\end{subequations}

\begin{align}
	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update}  
\end{align}
and	
\begin{align}
	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} 
\end{align}

where

\begin{align}
\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \quad \text{and}\\
\beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
\end{align}

\textcolor{red}{Next, we sample $\pi$ from its full conditional, given by}
%Next, the full conditional for $\pi$ is given by
\begin{align}
\label{eqn:pi}
	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

Lastly, we sample $\bm{Z}$ componentwise from the full conditional for each $Z_j$:
\begin{align}
	\label{eqn:z_full_conditional}
	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)},  \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
		1 - \pi^{(s+1)}, & q  = n_A + j, \\
	\end{cases}
\end{align}

where, for all $i \in \{1, \ldots, n_A\}$ and $j \in \{1, \ldots, n_B\}$, 
\begin{align}
	\label{eqn:fs_weight}
	w_{ij}^{(s)} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}^{(s)}}{u_{fl}^{(s)}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
\end{align}

\textcolor{teal}{
The full conditional in (\ref{eqn:z_full_conditional}) does not depend on $\bm{Z}^{(s)},$ implying that 
\begin{align}
p(\bm{Z}^{(s+1)} \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}) \propto \prod_q \left(\prod_j
p\left(Z_j^{(s+1)}  = q \mid \gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right)
\right).
\end{align}
}

\textcolor{teal}{Thus, we can make parallel updates to $\bm{Z}$ instead of sequential updates as in \cite{sadinle_bayesian_2017}.
In fact, to our knowledge, our approach and that of \cite{wortman2019} are the only two Bayesian approaches that propose parallel updates to the matching label (or linkage structure), \textcolor{red}{in contrast to} \textcolor{red}{other} Bayesian record linkage \textcolor{red}{methods} \citep{tancredi2011hierarchical, sadinle_detecting_2014, steorts_bayesian_2016, marchant_distributed_2019, zanella_flexible_2016, betancourt2022prior, aleshinguendel2021multifile}. }
%See Appendix~\ref{app:derivations} for derivations of the full conditional distributions.}

%Derivations for the full conditional distributions can be found in Appendix~\ref{app:derivations}.

\textcolor{teal}{Finally, to obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}. Since (\ref{eqn:fast_beta_prior}) does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$. To obtain a Bayes estimate that fulfills the \textcolor{red}{bipartite} requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. See Appendix \ref{bayes-estimate} for details regarding the initial Bayes estimate and \textcolor{red}{this} post-processing procedure.}

%Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.

\section{Efficient and Scalable Implementation}
\label{sec:efficiency}

The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A  n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors. One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}. In deterministic blocking, the modeler chooses fields thought to be highly reliable and only compares records that agree on those fields. The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains. Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
 undesirable %in some situations 
\citep{steorts_comparison_2014}.

After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields. However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the \textcolor{red}{classical} Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.

With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.

\hypertarget{data-representation-hashing-and-storage}{%
	\subsection{Data Representation, Hashing, and
		Storage}\label{data-representation-hashing-and-storage}}

Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
\textcolor{red}{The quantity $P$ is
 bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1),$ which clearly depends only on $F$ and $L_f.$ It does not depend (or scale with) with $n_A$ or $n_B.$}
% This comment below seems obvious and not needed.  
% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values.

%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.

>>>>

%%% only for appendix
\renewcommand{\appendixname}{Supplement}
\renewcommand{\thefigure}{SI\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{SI\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{SI\arabic{table}} \setcounter{equation}{0}
%%%

@@

%% *** Frontmatter *** 
\linenumbers
\begin{frontmatter}
\title{Supplementary Material for ``Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)"}
%\title{A Sample Document\thanksref{T1}}
%\thankstext{T1}{<thanks text>}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

@@

@@
